---
title: "machine_learning_project"
author: "Yucheng Xie"
date: "1/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load files and set seeds


```{r}
library("caret")
training <- read.csv("D:/pml-training.csv")
testing <- read.csv("D://pml-testing.csv")
set.seed(33833)
```

## preproccess--clean dataset
### drop first veariable and factorize user_name

```{r}
training <- training[,-1]
testing <- testing[,-1]
training$user_name <- as.factor(training$user_name)
testing$user_name <- as.factor(testing$user_name)
```

### change format to numeric and standardize all numeric variables
```{r}
v <- as.vector(1:159)
v <- v[-c(1,4,5,159)]
for (i in v){
  training[,i] <- as.numeric(as.character(training[,i]))
  training[,i] <- (training[,i] - mean(training[,i]))/sd(training[,i])
  testing[,i] <- as.numeric(as.character(testing[,i]))
  testing[,i] <- (testing[,i] - mean(testing[,i]))/sd(testing[,i])
}
```

### drop variables with more than 10000 na value
```{r}
na <- NULL
for (i in 1:159){
  nona <- sum(is.na(training[i]))
  na[i] <- nona
}
keep_variable <- na < 10000
training <- training[keep_variable]
testing <- testing[keep_variable]
```

### change time variable to factors indicating time(hour) within a day
I do it because the time that people do excercises may affect the quality of doing excersices
two other variables which is raw data for cvtd_timestamp varaible are deleted

```{r}
training$cvtd_timestamp <- sub(pattern = "[0-9]{2}/[0-9]{2}/[0-9]{4} ", replacement = "",training$cvtd_timestamp)
training$cvtd_timestamp <-sub(pattern = ":[0-9]{2}", replacement = "",training$cvtd_timestamp)
training$cvtd_timestamp <- as.factor(training$cvtd_timestamp)
training <- training[,-c(2,3)]

testing$cvtd_timestamp <- sub(pattern = "[0-9]{2}/[0-9]{2}/[0-9]{4} ", replacement = "",testing$cvtd_timestamp)
testing$cvtd_timestamp <-sub(pattern = ":[0-9]{2}", replacement = "",testing$cvtd_timestamp)
testing$cvtd_timestamp <- as.factor(testing$cvtd_timestamp)
testing <- testing[,-c(2,3)]
```

### change new_window into a dummy variable
```{r}
training[,3] <- as.numeric(training[,3])-1
testing[,3] <- as.numeric(testing[,3])-1
```

## preproccess--PCA analysis
I do principle component analysis so decrease the number of variables. Otherwise the machine-learning programm will take too much resourses
```{R}
prepro <- preProcess(training[,-c(1,2,57)],method = "pca",thresh = 0.8)
trainingpca <- predict(prepro, training[,-c(1,2,57)])
trainingpca <- (cbind(trainingpca,user_name = training[,1], timestamp = training$cvtd_timestamp, classe = training$classe))

testingpca <- predict(prepro, testing[,-c(1,2,57)])
testingpca <- (cbind(testingpca,user_name = testing[,1], timestamp = testing$cvtd_timestamp))
```

## create folds for cross validation
create 5 folds, use 4 folds as training set, 1 folds as valid set. do it 5 times 
```{R}
folds <- createFolds(y=trainingpca$classe,k=5,list=FALSE)
trainingpca <- cbind(trainingpca,folds)
for (i in 1:5){
  traininglist <- split(trainingpca,trainingpca$folds)
  assign(paste("testing", i, sep = ""),(traininglist[[i]][,-18]))
}
training1 <- rbind(testing2,testing3,testing4,testing5)
training2 <- rbind(testing3,testing4,testing5,testing1)
training3 <- rbind(testing4,testing5,testing1,testing2)
training4 <- rbind(testing5,testing1,testing2,testing3)
training5 <- rbind(testing1,testing2,testing3,testing4)  
```

## do random forest analysis within each fold
I only use random forest methods because other mothods (boosting, generalized linear model) get less accurate results. The same for assemled models.
```{R}
mod1rf <- train(classe~.,data = training1, method = "rf")
mod2rf <- train(classe~.,data = training2, method = "rf")
mod3rf <- train(classe~.,data = training3, method = "rf")
mod4rf <- train(classe~.,data = training4, method = "rf")
mod5rf <- train(classe~.,data = training5, method = "rf")
```

## predict data in valid sets and get accuracy; set accuracy rate as accu_weigh
I input the accuracy be myself
```{R}
pred1rf <- predict(mod1rf,newdata = testing1)
pred2rf <- predict(mod2rf,newdata = testing2)
pred3rf <- predict(mod3rf,newdata = testing3)
pred4rf <- predict(mod4rf,newdata = testing4)
pred5rf <- predict(mod5rf,newdata = testing5)
accu1 <- confusionMatrix(pred1rf,testing1$classe)
accu2 <- confusionMatrix(pred2rf,testing2$classe)
accu3 <- confusionMatrix(pred3rf,testing3$classe)
accu4 <- confusionMatrix(pred4rf,testing4$classe)
accu5 <- confusionMatrix(pred5rf,testing5$classe)
accu1$
accu_weigh1 <- 0.9682
accu_weigh2 <- 0.9694
accu_weigh3 <- 0.9673
accu_weigh4 <- 0.9643
accu_weigh5 <- 0.9658
```

## predict the results in testing set with 5 models
```{R}
pred1test <- predict(mod1rf,newdata = testingpca)
pred2test <- predict(mod1rf,newdata = testingpca)
pred3test <- predict(mod1rf,newdata = testingpca)
pred4test <- predict(mod1rf,newdata = testingpca)
pred5test <- predict(mod1rf,newdata = testingpca)
predmatrix <- cbind(pred1test,pred2test,pred3test,pred4test,pred5test)
```

## vote (weighed by accuracy) to get final results
weigh equals to accuracy of each model get from previous steps
```{R}
add_weigh <- function(factor, pred1, pred2, pred3, pred4, pred5 ){
  weigh <- 0
  if (pred1 == factor){weigh <- weigh + accu_weigh1}
  if (pred2 == factor){weigh <- weigh + accu_weigh2}
  if (pred3 == factor){weigh <- weigh + accu_weigh3}
  if (pred4 == factor){weigh <- weigh + accu_weigh4}
  if (pred5 == factor){weigh <- weigh + accu_weigh5}
  return(weigh)
}

vote <- function(pred1, pred2, pred3, pred4, pred5){
  weigha <- add_weigh(1 ,pred1, pred2, pred3, pred4, pred5)
  weighb <- add_weigh(2 ,pred1, pred2, pred3, pred4, pred5)
  weighc <- add_weigh(3 ,pred1, pred2, pred3, pred4, pred5)
  weighd <- add_weigh(4 ,pred1, pred2, pred3, pred4, pred5)
  weighe <- add_weigh(5 ,pred1, pred2, pred3, pred4, pred5)
  if (all(weigha>weighb, weigha>weighc, weigha>weighd, weigha>weighe)){return("a")}
  else if (all(weighb>weighc, weighb>weighd, weighb>weighe)){return("b")}
  else if (all(weighc>weighd,weighc>weighe)){return("c")}
  else if (all(weighd>weighe)){return("d")}
  else{return("e")}
}
result <- NULL
for (i in 1:nrow(testing)){
  resulti <- vote(predmatrix[i,1],predmatrix[i,2],predmatrix[i,3],predmatrix[i,4],predmatrix[i,5]) 
}
```

##possible out-of sample errors
according to the cross validation, the out of sample error is approimately 5% because the accuracy is about 95%
Also, in my codes there are many repeating sentences. Do you know how to make them shorter (maybe with loop or something). Thank you








